Constraining gravitational wave amplitude birefringence with GWTC-3

### 2022-12-10 src/tex/ms.tex
Formatting
Update content

With the posterior of $\kappa$ from PE on each event, the posterior of $\mu$ and $\sigma$ can be calculated by
\begin{equation}
    p(\mu,\sigma|\{d\})\propto p(\mu,\sigma)\prod_{i}\int\frac{p(\kappa_i|d_i)}{p(\kappa_i}p(\kappa_i|\mu,\sigma)\,d\kappa_i\,,
    \label{eq:posterior_of_mu_sigma}
\end{equation}
Note that $p(\kappa)$ is the prior of $\kappa$, which is a uniform distribution between $-1$ and $1$ in this study.
We chose the prior of $\mu$ and $\sigma$ being uniform distributions.
Equation \ref{eq:posterior_of_mu_sigma} can then be simplified to
\begin{equation}
    p(\mu,\sigma|\{d\})\propto\prod_{i}\int p(\kappa_i|d_i)\mathcal{N}(\mu,\sigma)\,d\kappa_i\,.
\end{equation}
To sample the posterior distribution of $\mu$ and $\sigma$, we used a sampling package, flowMC \citep{flowMC}.

@@

\subsection{Bayesian hierarchical modeling}
% Corner plot of the mean and standard deviation of kappa
The posterior of $\mu$ and $\sigma$ are shown in figure \ref{fig:corner_Gaussian}.
The median of $\mu$ within $90\%$ confidence interval is \variable{output/mu_median.txt}.
The median of $\sigma$ within $90\%$ confidence interval is \variable{output/sigma_median.txt}.
The posterior of $\mu$ is concentrated around $0$, which means the population constraint on $\kappa$ is consistent with GR.

@@

\begin{figure}[h]
    \script{posterior_kappa.py}
    \includegraphics[width=\columnwidth]{figures/posterior_kappa.pdf}
    \caption{
        The generic and restricted posterior of $\kappa$.
        The orange solid line shows the restricted posterior of $\kappa$, and the orange dashed lines enclose the $90\%$ confidence interval.
        The blue solid line shows the generic posterior of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval.
        The pink vertical solid line represents $\kappa=0$.
    }
    \label{fig:posterior_kappa}
\end{figure}

@@

% Comparison with previous studies
\subsection{Comparison with Previous Studies}
\citet{Okounkova_2022} gave a constraint on GW amplitude birefringence by performing statistical analysis on GWTC-2.
We convert the constraint to the same format as ours to make a comparison.
They were able to constrain $\kappa$ to be $|\kappa| \lesssim 0.74$ at $1 \sigma$.
We obtained a tighter constraint on $\kappa$ with $|\kappa| \lesssim 0.04$ at $1 \sigma$.
Our result is an order of magnitude improvement in constraining $\kappa$.
The main reason is that we have more events from GWTC-3 than GWTC-2.

% Future work
\subsection{Future Work}
% BNS
Future work is to perform PE on binary neutron star mergers, such as GW170817.
The frequency range of the signal is much wider compared to the binary black hole mergers.
Thus, the difference in the effect of birefringence at different frequencies within the range can be more significant.
The PE on binary neutron star mergers can allow us to further constrain the birefringence effect and the beyond-GR theories that predict it.
However, performing PE on binary neutron star mergers requires much more computational resources.
Therefore, we may need to wait for future PE methods and tools to further our work.

@@

\section{Acknowledgements}
\label{sec:Acknowledgements}
M.~I., K.~W.~K.~W.~, and W.~M.~F.~ are funded by the Center for Computational Astrophysics at the Flatiron Institute.
The Flatiron Institute provided the computational resources used in this work.

This research has made use of data or software obtained from the Gravitational Wave Open Science Center (gwosc.org), a service of LIGO Laboratory, the LIGO Scientific Collaboration, the Virgo Collaboration, and KAGRA.
LIGO Laboratory and Advanced LIGO are funded by the United States National Science Foundation (NSF) as well as the Science and Technology Facilities Council (STFC) of the United Kingdom, the Max-Planck-Society (MPS), and the State of Niedersachsen/Germany for support of the construction of Advanced LIGO and construction and operation of the GEO600 detector.
Additional support for Advanced LIGO was provided by the Australian Research Council.
Virgo is funded, through the European Gravitational Observatory (EGO), by the French Centre National de Recherche Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) and the Dutch Nikhef, with contributions by institutions from Belgium, Germany, Greece, Hungary, Ireland, Japan, Monaco, Poland, Portugal, Spain.
KAGRA is supported by Ministry of Education, Culture, Sports, Science and Technology (MEXT), Japan Society for the Promotion of Science (JSPS) in Japan; National Research Foundation (NRF) and Ministry of Science and ICT (MSIT) in Korea; Academia Sinica (AS) and National Science and Technology Council (NSTC) in Taiwan.

% \pagebreak
\bibliography{bib}

>>>>

With the posterior of $\kappa$ from PE on each event, the posterior of $\mu$ and $\sigma$ can be calculated by
\begin{equation}
    p(\mu,\sigma|\{d\})\propto p(\mu,\sigma)\prod_{i}\int\frac{p(\kappa_i|d_i)}{p(\kappa_i)}p(\kappa_i|\mu,\sigma)\,d\kappa_i\,,
    \label{eq:posterior_of_mu_sigma}
\end{equation}
Note that $p(\kappa)$ is the prior of $\kappa$, which is a uniform distribution between $-1$ and $1$ in this study.
We chose the prior of $\mu$ and $\sigma$ being uniform distributions.
Equation \ref{eq:posterior_of_mu_sigma} can then be simplified to
\begin{equation}
    p(\mu,\sigma|\{d\})\propto\prod_{i}\int p(\kappa_i|d_i)\mathcal{N}(\mu,\sigma)\,d\kappa_i\,.
\end{equation}
To sample the posterior distribution of $\mu$ and $\sigma$, we used a sampling package, flowMC \citep{flowMC}.

@@

\subsection{Bayesian hierarchical modeling}
% Corner plot of the mean and standard deviation of kappa
We then show the result of Bayesian hierarchical modeling in figure \ref{fig:corner_Gaussian}.
The median of $\mu$ within $90\%$ confidence interval is \variable{output/mu_median.txt}.
The median of $\sigma$ within $90\%$ confidence interval is \variable{output/sigma_median.txt}.
The posterior of $\mu$ is concentrated around $0$, which means the population constraint on $\kappa$ is consistent with GR.

@@

\begin{figure}[h]
    \script{posterior_kappa.py}
    \includegraphics[width=\columnwidth]{figures/posterior_kappa.pdf}
    \caption{
        The generic and restricted posterior of $\kappa$.
        The blue solid line shows the restricted posterior of $\kappa$, and the blue dashed lines enclose the $90\%$ confidence interval.
        The orange solid line shows the generic posterior of $\kappa$, and the orange dashed lines enclose the $90\%$ confidence interval.
        The pink vertical solid line represents $\kappa=0$.
    }
    \label{fig:posterior_kappa}
\end{figure}

@@

% Comparison with previous studies
\subsection{Comparison with previous studies}
\citet{Okounkova_2022} gave a constraint on GW amplitude birefringence by performing statistical analysis on GWTC-2.
We convert the constraint to the same format as ours to make a comparison.
They were able to constrain $\kappa$ to be $|\kappa| \lesssim 0.74$ at $1 \sigma$.
We obtained a tighter constraint on $\kappa$ with $|\kappa| \lesssim 0.04$ at $1 \sigma$.
Our result is an order of magnitude improvement in constraining $\kappa$.
The main reason is that we have more events from GWTC-3 than GWTC-2.

% Future work
\subsection{Future work}
% BNS
Future work is to perform PE on binary neutron star mergers, such as GW170817.
The frequency range of the signal is much wider compared to the binary black hole mergers.
Thus, the difference in the effect of birefringence at different frequencies within the range can be more significant.
The PE on binary neutron star mergers can allow us to further constrain the birefringence effect and the beyond-GR theories that predict it.
However, performing PE on binary neutron star mergers requires much more computational resources.
Therefore, we may need to wait for future PE methods and tools to further our work.

@@

\begin{acknowledgments}
M.~I., K.~W.~K.~W.~, and W.~M.~F.~ are funded by the Center for Computational Astrophysics at the Flatiron Institute.
The Flatiron Institute provided the computational resources used in this work.

This research has made use of data or software obtained from the Gravitational Wave Open Science Center (gwosc.org), a service of LIGO Laboratory, the LIGO Scientific Collaboration, the Virgo Collaboration, and KAGRA.
LIGO Laboratory and Advanced LIGO are funded by the United States National Science Foundation (NSF) as well as the Science and Technology Facilities Council (STFC) of the United Kingdom, the Max-Planck-Society (MPS), and the State of Niedersachsen/Germany for support of the construction of Advanced LIGO and construction and operation of the GEO600 detector.
Additional support for Advanced LIGO was provided by the Australian Research Council.
Virgo is funded, through the European Gravitational Observatory (EGO), by the French Centre National de Recherche Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) and the Dutch Nikhef, with contributions by institutions from Belgium, Germany, Greece, Hungary, Ireland, Japan, Monaco, Poland, Portugal, Spain.
KAGRA is supported by Ministry of Education, Culture, Sports, Science and Technology (MEXT), Japan Society for the Promotion of Science (JSPS) in Japan; National Research Foundation (NRF) and Ministry of Science and ICT (MSIT) in Korea; Academia Sinica (AS) and National Science and Technology Council (NSTC) in Taiwan.
\end{acknowledgments}

\bibliography{bib}
